---
title: "R Notebook"
output: html_notebook
---
### Assigment 1
Preparations!
```{r}
s = 14
f = 6
a0 = 2
b0 = 2
```
#### 1a
We use the rbeta function to draw from the posterior distribution. We do it for different sample sizes for visualization later.
```{r}
d10 = rbeta(10,a0+s,b0+f)
d100 = rbeta(100,a0+s,b0+f)
d1000 = rbeta(1000,a0+s,b0+f)
```
Plot the result, as number of draws increases, the curve looks more like the posterior.
```{r}
x = 1:10/10
y_max = max(max(d10),max(d100),max(d1000),4)+1
plot(density(d10),ylim=c(0,y_max),xlim=c(0,1), col = "red",xlab="",main = sprintf("Draws = {10,100,1000}"))
lines (density(d100),col = "blue")
lines (density(d1000),col = "green")
curve(dbeta(x,2+14,2+6),add = TRUE,col = "yellow")
legend(0,y_max, legend=c("draws = 10", "draws = 100","draws = 1000","Beta"),
       col=c("red", "blue","green","yellow"), lty=1:2, cex=0.8)
```
#### 1b
Draw from the posterior using rbeta and just count the values that are less then the given threshold.
```{r}
draws = rbeta(10000,a0+s,b0+f)
dprob = NROW(draws[draws < 0.4])/NROW(draws)
bprob = pbeta(0.4,a0+s,b0+f)
print(sprintf("Draw-prob = %f and Beta-prob = %f",dprob,bprob))
```
#### 1c
Use the draw results from 1b and transform them into log-odds.
```{r}
log_odds = log(draws/(1-draws))
plot(hist(log_odds,breaks=50))
```

### Assigment 2
Preparations!
```{r}
u = 3.5
income = c(14,25,45,25,30,33,19,50,34,67)
tau2 = sum(log(income)-u)^2/length(income)
```
#### 2a
The geoR::rinvchisq function simulates draws from scaled inverse chi-square distribution. From the assigment, we have two parameters n and $$\tau^2$$ represented as scale and df (degree of freedom) paramaters to the geoR::rinvchisq function. Generate draws and plot it with the density function (geoR::dinvchisq).
```{r}
draws = geoR::rinvchisq(10000,df=length(income),scale=tau2);

# Offset not defined for 0
x = 0:1000/1000+1/100000;
h = hist(draws,breaks=50)
d = geoR::dinvchisq(x,df=length(income),scale=tau2)
ymax = max(h$density,d)+1;
plot(h,ylim=c(0,ymax),freq=FALSE)
lines(x,d)
```
#### 2b
The build in pnorm function gives us the CDF that is normally distributed. Using the results from 2a, we can plot the values from the Gini coefficient.
```{r}
g = 2*pnorm(sqrt(draws)/sqrt(2))-1
hist(g,breaks = 50)
```
#### 2c
Take out 2.5% from each tail in the distribution and density function, plot them to see visually the difference.
```{r}
g = g[order(g)];
g_start = floor(length(g)*0.025)
g_stop = floor(length(g)*0.975)
density_g = density(g)
gd_start = floor(length(density_g$x)*0.025)
gd_stop = floor(length(density_g$x)*0.975)
plot(density_g$x[gd_start:gd_stop],density_g$y[gd_start:gd_stop],type="l")
hist(g[g_start:g_stop],add=TRUE,freq=FALSE,breaks =30)
```
### Assigment 3

